Misperceived Social Norms about Women Working Outside the Home and Its Effect on Women in Saudi Arabia

### 2024-02-14 outputs/paper/paper.qmd
modify abstract

---
title: "Misperceived Social Norms about Women Working Outside the Home and Its Effect on Women in Saudi Arabia"
author: 
  - Boxuan Yi
  - Ruoxian Wu
thanks: "Code and data are available at: https://github.com/ScarletWu/Misperceived-Social-Norms-Women-Working-Outside-the-Home-in-Saudi-Arabia. Replication on Social Science Reproduction platform is available at: https://www.socialsciencereproduction.org/reproductions/1557/"
date: "February 12, 2024"
date-format: long
abstract: "Husbands who hold guardianship and discourage their wives from working outside the home are often perceived as the main reason for the low female workforce participation in Saudi Arabia. Bursztyn, González, and Yanagizawa-Drott explore this social norm through surveys, experiments and visualizations, revealing that a majority of young husbands in Saudi Arabia privately support women working outside the home but underestimate the support of other men, and correcting this misperception increases husbands’ willingness to help wives search for jobs. This paper replicates the data from their study to analyze young husbands’ actual opinions and perceptions, confirming similar findings. This paper also discusses strategies for reducing misperceptions and the unusual Saudi female employment rate during the pandemic."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Introduction
Barriers to the employment of women exert a significant influence on female labor, presenting a constraint to increasing economic opportunities for women [@sciencedirectOrganizationalEconomic]. Saudi Arabia is a recognizable example of a country with a low rate of female labor force participation. In 2022, the female labor force participation rate, defined as the percentage of the female population aged 15 and above that is employed, is only 27.8% in Saudi Arabia. This is notably lower than the global average of 47.3%, while the male labor force participation rate in Saudi Arabia remains high at 80% [@worldbankLaborForce].

@@

@@

The dataset resulting from online survey comprises 1460 observations and 21 variables. These variables include condition, age, employed_now, employed_wife, employed_out_wife, children, and the responses to the survey questions for both groups as indicated above, along with participants’ confidence levels in some of the answers. Details and explanations using visualization will be further discusses in the next section.

@@

@fig-job_sign_up displays the main result from the experimental intervention, illustrating the difference in job matching service sign-up rates between the control and treatment groups. In the control group, 23.48% of participants are willing to sign up their wives for the job service when provided with another gift card option. In the treatment group, where participants are given the actual proportion of support for WWOH, 32.02% of the respondents prefer to sign up their wives for the job service, which is a 32% increase (with a p-value of 0.0424).

The combination of @tbl-summary_table and @fig-job_sign_up demonstrates the effects of the information intervention on participants’ behavior and opinions towards women's labor market participation, revealing not only a statistical significance but also practical importance in the context of social norms and labor supply outcomes in Saudi Arabia.

\newpage
The longer-term outcomes of the experiment on information intervention are shown in @fig-3 with error bars indicating the 95% confidence intervals for these proportions.

The percentage of wives applying for a job outside the home is presented in @fig-3 a. In the control group, 6.04% of wives apply for jobs outside, while 16.74% wives in the treatment group choose to do so. The increase in application rate is quite obvious, nearly a 180% increase with a p-value of 0.001. @fig-3 b shows the impact of the informational intervention on outside-home job interviews. The interview rate in the control group is 1.1% while the treatment group has 5.95%, with p-value of 0.013. Both @fig-3 a and @fig-3 b display significant increases.

Regarding the employment rates for this kind of job, @fig-3 c shows that the employment rate increases from 7.69% to 9.73%. Although this indicates an increase, it is not statistically significant (p-value = 0.456), implying that though the intervention might have had an encouraging effect on job-seeking behavior, it does not translate into a significant difference in actual employment within the study period.

To further explore the long-term influence, @fig-3 d shows the share of husbands who report in a hypothetical question whether they would sign their wives up for driving lessons. The control group has a rate of 68.13%, and the treatment group has a rate of 78.38%, with p-value as 0.03. The increase in husbands’ willingness to sign their wives up for driving lessons implies that correcting misperception influences beyond behaviours associated with female employment, creating potential impacts on broader social norms and attitudes toward women's conditions in Saudi Arabia.

Overall, the treatment has a significant effects on both immediate and longer-term outcomes related to women labor market participation based on @fig-3.

@@

@fig-national_survey_misperception is created to illustrate the results of the national survey conducted with 1460 Saudi men aged from 18 to 35. The x-axis represents the share of others agreeing, and the y-axis denotes the cumulative probability that participants agree with the pro-WWOH statement. Two vertical lines in @fig-national_survey_misperception represent the actual proportion of respondents who agree with the statement.

There is no significant difference between the data concerning the control group and the treatment group. @fig-national_survey_misperception reveals that the average proportion of support for both groups is slightly above 80%, while more than 90% of men underestimate this proportion. The number of underestimations is larger than the number observed in the main experiment, and one explanation is that participants are no longer being asked about opinions from people that they are familiar.

>>>>

---
title: "Misperceived Social Norms about Women Working Outside the Home and Its Effect on Women in Saudi Arabia"
author: 
  - Boxuan Yi
  - Ruoxian Wu
thanks: "Code and data are available at: https://github.com/ScarletWu/Misperceived-Social-Norms-Women-Working-Outside-the-Home-in-Saudi-Arabia. Replication on Social Science Reproduction platform is available at: https://www.socialsciencereproduction.org/reproductions/1557/"
date: "February 12, 2024"
date-format: long
abstract: "Husbands who hold guardianship and discourage their wives from working outside the home are often perceived as the main reason for the low female workforce participation in Saudi Arabia. We use data and surveys from the American Economic Association to analyze this perception, revealing that a majority of young husbands in Saudi Arabia privately support women working outside the home but underestimate the support of other men. This paper also discusses the positive effects brought about by correcting this misperception, strategies for reducing information intervention, and the unusual Saudi female employment rate during the pandemic."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Introduction
Barriers to the employment of women exert a significant influence on female labor, presenting a constraint to increasing economic opportunities for women [@sciencedirectOrganizationalEconomic]. Saudi Arabia is an example of a country with a low rate of female labor force participation. In 2022, the female labor force participation rate, defined as the percentage of the female population aged 15 and above that is employed, is only 27.8% in Saudi Arabia. This is notably lower than the global average of 47.3%, while the male labor force participation rate in Saudi Arabia remains high at 80% [@worldbankLaborForce].

@@

The variables from the main experiment dataset that we will use are condition2 (control or treatment group), age, children (number of children), college_deg (whether the participant has college degree or above), employed_now (if the participant is employed when conducting the experiment), employed_wife (if the participant’s wife is employed when conducting the experiment), num_know_per (percentage of people known in the session), num_mfs_per (percentage of people in the session with mutual friends), signed_up_number (whether the husband sign wife up for job matching), outside_wedge (guessing number of people in the session supporting WWOH minus the actual number of people in the session supporting WWOH). Variables will be used in the follow-up dataset are condition2 (control or treatment group), employed_3mos_out_fl (if the participant’s wife was employed three months ago when conducting the experiment), applied_out_fl (if the wife apply for job working outside), interviewed_out_fl (if the wife interview for job working outside), employed_now_out_fl (if the wife be employed for job outside), and driving_fl (if the husband is willing to sign his wife up for driving lessons).

@@

The dataset resulting from online survey comprises 1460 observations and 21 variables. These variables include condition, age, employed_now, employed_wife, employed_out_wife, children, and the responses to the survey questions for both groups as indicated above, along with participants’ confidence levels in some of the answers. Details and explanations using visualization will be further discusses in the next section. The variables will be used are c_outside_guess_frac (guessing the fraction that support WWOH in the control group), t_outside_guess_frac (guessing the fraction that support WWOH in the treatment group), c_outside_mean (the mean of actual proportion of support in the control group), t_outside_mean (the mean of actual proportion of support in the treatment group).

@@

@fig-job_sign_up displays the main result from the experimental intervention, illustrating the difference in job matching service sign-up rates between the control and treatment groups. In the control group, 23.48% of participants are willing to sign up their wives for the job service when provided with another gift card option. In the treatment group, where participants are given the actual proportion of support for WWOH, 32.02% of the respondents prefer to sign up their wives for the job service, which is a 32% increase (with a p-value of 0.0424). The combination of @tbl-summary_table and @fig-job_sign_up demonstrates the effects of the information intervention on participants’ behavior and opinions towards women's labor market participation, revealing not only a statistical significance but also practical importance in the context of social norms and labor supply outcomes in Saudi Arabia.

\newpage
The longer-term outcomes of the experiment on information intervention are shown in @fig-3 with error bars indicating the 95% confidence intervals for these proportions. The percentage of wives applying for a job outside the home is presented in @fig-3 a. In the control group, 6.04% of wives apply for jobs outside, while 16.74% wives in the treatment group choose to do so. The increase in application rate is quite obvious, nearly a 180% increase with a p-value of 0.001. @fig-3 b shows the impact of the informational intervention on outside-home job interviews. The interview rate in the control group is 1.1% while the treatment group has 5.95%, with p-value of 0.013. Both @fig-3 a and @fig-3 b display significant increases. Regarding the employment rates for this kind of job, @fig-3 c shows that the employment rate increases from 7.69% to 9.73%. Although this indicates an increase, it is not statistically significant (p-value = 0.456), implying that though the intervention might have had an encouraging effect on job-seeking behavior, it does not translate into a significant difference in actual employment within the study period.

To further explore the long-term influence, @fig-3 d shows the share of husbands who report in a hypothetical question whether they would sign their wives up for driving lessons. The control group has a rate of 68.13%, and the treatment group has a rate of 78.38%, with p-value as 0.03. The increase in husbands’ willingness to sign their wives up for driving lessons implies that correcting misperception influences beyond behaviours associated with female employment, creating potential impacts on broader social norms and attitudes toward women's conditions in Saudi Arabia. Overall, the treatment has a significant effects on both immediate and longer-term outcomes related to women labor market participation based on @fig-3.

@@

@fig-national_survey_misperception is created to illustrate the results of the national survey conducted with 1460 Saudi men aged from 18 to 35. The x-axis represents the share of others agreeing, and the y-axis denotes the cumulative probability that participants agree with the pro-WWOH statement. Two vertical lines in @fig-national_survey_misperception represent the actual proportion of respondents who agree with the statement. There is no significant difference between the data concerning the control group and the treatment group. @fig-national_survey_misperception reveals that the average proportion of support for both groups is slightly above 80%, while more than 90% of men underestimate this proportion. The number of underestimations is larger than the number observed in the main experiment, and one explanation is that participants are no longer being asked about opinions from people that they are familiar.
